Forecasting future oil production in Norway and the UK: a general improved methodology

### 2014-06-23 publications/Forecasting  Future Oil Production/peak_oil_4.tex
Latest version.

The idea behind the methodology is to model the future aggregate oil
production of a country by studying the production dynamics of its
individual constituents, the oil fields. The main benefit of this
approach, compared to working directly with aggregate production data,
is the possibility to forecast non-trivial oil production profiles
arising from the combination of all the individual field\textquoteright{}s
dynamics. In order to achieve that, one must be able 1) to \textbf{extend
the oil production of each individual field} into the future and 2)
to \textbf{extrapolate the rate of discoveries} of new oil fields.

@@

The first step to predicting the future oil production of a country
is to extrapolate the future production of existing fields and to
estimate the error on this extrapolation. The data of the fields developed
in the past shows a repeating asymmetric pattern. A good example is
the Oseberg field shown in figure \ref{regular}, with a quick ramp
up once the field is being developed, and then a peak or plateau before
the field starts decaying. The decay can take many different shapes
and is governed by a variety of geological and economical factors
which will be captured implicitly by the fitting procedure.

@@

To determine how well the extrapolation based on the stretched exponential
predicts the future production, a complete back-testing has been performed
on each field. A single back-test is made as follows:
\begin{itemize}
\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field is
truncated at a certain date in the past $\tau\in\{0,\,\ldots,\, N\}$.
Where $\tau$ is in months since production start of the field.
\item The extrapolation$f(t)$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{\tau}\}$.
\item The future production predicted by the extrapolation $f(t)$ can be
compared to the actual production from the date $\tau$ in the past
up to now $\tau_{f}=N$. The extrapolated total production can be
computed as 
\begin{equation}
P_{e}(\tau)=\int_{\tau}^{\tau_{f}}f(t)dt
\end{equation}
and the relative error is given by
\begin{equation}
e(\tau)=\frac{P_{e}(\tau)-\sum_{i=\tau}^{\tau_{f}}p_{i}}{P_{e}(\tau)},
\end{equation}
where both are a function of the truncation time $\tau$.
\end{itemize}
Computing this back test for every month in the past since the field
starting decaying, yields a plot showing the evolution of the relative
error over time \ref{oesberg-error}. By construction, the relative
error will tend to zero as the truncation time $\tau$ goes to the
present. Nonetheless, it is a uself indicator for the stability of
the extrapolation. As can been seen for the Oseberg field, the relative
error on future production remained fairly stable during the past
decade.

From the complete back-test, on can compute the average relative error
\begin{equation}
\bar{e}=\frac{1}{N}\sum_{i=0}^{N}e(i)
\end{equation}
 the extrapolation made on the future production. Assuming that the
relative errors are normally distributed around the average relative
error, the standard deviation on the average relative error is given
by 
\begin{equation}
\sigma_{e}=\sqrt{\frac{1}{N}\sum_{i=0}^{N}\left(e(i)-\bar{e}\right)^{2}}.\label{eq:field-risk}
\end{equation}
As the average relative error is often fairly constant, the extrapolation
was corrected by the average relative error. Meaning that if the extrapolation
consistently over-estimated the production by 10\% during the back-test,
the extrapolation was reduced by 10\%. This results in an extrapolated
production $p(t)$, including a 1$\sigma$ confidence interval, as
\begin{equation}
p(t)=(1-\bar{e})\, f(t)\pm\sigma_{e}.\label{eq:extrapolation}
\end{equation}

An example of such an extrapolation including a one standard deviation
range is shown on figure \ref{irregular} for the Oesberg field.

@@

Once the individual fields have been extrapolated in the sense of
formula \ref{eq:extrapolation}, the aim is to compute the extrapolation
for the country. While it is straightforward to sum the extrapolations
of the individual fields to obtain the expected production, some care
has to be taken with respect to the confidence interval on the production
at the country level.

As shown later in section \ref{sec:Results}, the same extrapolation
including a complete monthly back-test of total future production
has been performed at the country level and the resulting relative
error is much smaller than the average error observed on the individual
fields. To account for this observation, the assumption made is that
the relative error between individual fields is uncorrelated. Therefore,
the fields can be considered as a portfolio of assets with a return
given by their extrapolation $p(t)$ (eq. \ref{eq:extrapolation})
and a risk given by $\sigma_{field}$ (eq. \ref{eq:field-risk}).
This means that the risk at the country level from the extrapolated
production is to be computed as 
\begin{equation}
\sigma_{country}^{2}=\frac{1}{\#fields}\sum_{field\in fields}\sigma_{field}^{2}.
\end{equation}

Intuitively, this models well the fact that the uncorrelated errors
among fields will mostly cancel out.

@@

Knowing the future production rate of existing fields is not enough
as new fields will be discovered in the future. The model describing
the discovery rate of new fields should satisfy two fundamental observations.
\begin{enumerate}
\item The rate of new discoveries should tend to zero as time goes to infinity.
This is a consequence of the finiteness of the number of oil fields. 
\item The rate of new discoveries should depend on the size of the oil fields.
As of today, giant oil fields are discovered much less frequently
than dwarfs.
\end{enumerate}

@@

A natural choice for such a model is a non-homogenous Poisson process.
The Poisson process is a process that generates independent events
at a rate $\lambda$. It is inhomogeneous if the rate is time-dependent,
$\lambda\rightarrow\lambda(t)$. The standard way to measure $\lambda(t)$
is to find a functional form for $N(t)$, the statistical average
of the cumulative number of events (discoveries) up to time $t$.
Then, $\lambda(t)$ is simply $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
shows $N(t)$ for Norwegian fields classified according to their size.
The logistic curve is a good fit to the data (integral form of equation
\ref{logistic}). This implies that after an initial increase, the
rate of new discoveries reaches a peak followed by a decrease until
no more oil fields are to be found, consistent with our fundamental
observations. This same approach has already been successfully applied
by Forro \cite{Forro 2012} to estimate the number of daily active
users on Zynga.

As the discovery and production dynamics are not independent of the
field size, the fields have been split into two groups: dwarfs and
giants. Unfortunately, the two logistic curves thus obtained are highly
sensitive to the splitting size. This results from the major issue,
when fitting a logistic curve to data, that the carrying capacity
can not be determined if the data does not already exhibit the slow
down in growth from reaching the carrying capacity. However, it is
mentioned in the literature that often dwarf fields have already been
discovered a long time ago, but their production has been postponed
for economical reasons \cite[p. 378]{Lynch 2002}. Therefore, it is
expected that the large oil fields have mostly been found and produced,
and that future discoveries will mostly be made up of dwarf fields.
Consequently, the splitting size has been picked as small as possible
in order to maximize the number of giant fields but large enough to
avoid recent discoveries:
\begin{itemize}
\item \textbf{Dwarfs}: Fields which produced less than $50\cdot10^{6}$
barrels.
\item \textbf{Giants}: Fields which produced more than $50\cdot10^{6}$
barrels.
\end{itemize}
The resulting plot shown in figure \ref{rate_of_discovery} pictures
the dynamics: giant oil fields have mostly been found while the discovery
process for dwarf fields is still ongoing. The logistic growth curve
fit to the giant discoveries is well constrained, however the fit
for the dwarfs is poorly constrained. As can be seen on figure \ref{rate_of_discovery},
the carrying capacity $K$ of the logistic growth model is not well
constrained by the available data. A large spectrum of values for
$K$ can lead to an equally good fit of the data. Therefore, a method
described in section \ref{sub:Likelihood-function-for} has been used
to compute the probability of different carrying capacities.

@@

To overcome the logistic fitting issue for the dwarf fields, a method
already used by Smith \cite{Smith 1980} has been implemented. This
method makes the following two postulates:
\begin{enumerate}
\item ``The discovery of reservoirs in a petroleum play can be modeled
statistically as sampling without replacement from the underlying
population of reservoirs.''
\item ``The discovery of a particular reservoir from among the existing
population is random, with a probability of discovery being dependent
on reservoir size.''
\end{enumerate}
The fields are split into $J$ size bins denoted $S_{1},\,\ldots,\, S_{J}$
occurring with frequency $n_{1},\,\ldots,\, n_{J}$. Each discovery
is considered as a step $i$ at which a field of size $I(i)\in\{S_{1},\,\ldots,\, S_{J}\}$
is found and $m_{ij}$ denotes the number of fields of size $j\in\{1,\,\ldots,\, J\}$
discovered before the $i^{th}$ step. Than the probability that the
discovery at step $i^{th}$ is of size $j$ can be expressed as
\begin{equation}
P\left(I(i)=S_{j}\right)=\frac{\left(n_{j}-m_{ij}\right)\cdot S_{j}}{\sum_{k=1}^{J}\left(n_{k}-m_{ik}\right)\cdot S_{k}}.
\end{equation}
The likelihood $L$ for a complete sequence of $N$ discoveries $\{I(1),\,\ldots,\, I(N)\}$
can than be expressed as
\begin{equation}
L=\prod_{i=1}^{N}\frac{\left(n_{I(i)}-m_{iI(i)}\right)\cdot S_{I(i)}}{\sum_{j=1}^{J}\left(n_{j}-m_{ij}\right)\cdot S_{j}}.\label{eq:likelihood}
\end{equation}
The unknown parameters are the number of fields $n_{1},\,\ldots,\, n_{J}$,
whose likelihood can now be estimated based on the existing discoveries.
Using a brute force approach, the entire space of plausible values
for the variables $n_{1},\,\ldots,\, n_{J}$ has been sampled. The
values $n_{j}$ have been sampled between the number of existing fields
$m_{Nj}$ in the bin $j$ and up to a value $n_{j}^{upper}$, such
that the scenario with the largest likelihood according to equation
\ref{eq:likelihood} has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
the bin $j$. Subsequently, the likelihood of each scenario (value
of the tuple $n_{1},\,\ldots,\, n_{J}$) has been normalized such
that the total likelihood of all generated scenarios equals one.

For the analysis of the north sea oil field discoveries, the number
of size bins has been fixed to $J=2$ splitting between dwarfs (1)
and giants (2) as described in section \ref{sub:Discoveries-modeled-as}.

\begin{table}[h]
\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.}

\centering %
\begin{tabular}{lllll}
 & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{likelihood-results}
\end{table}

The results shown in table \ref{likelihood-results} are coherent
with the intuitive expectation that discovering a new giant field
is unlikely and that future discoveries wil mostly be made up of dwarf
fields. The likelihoods obtained for the carrying capacities of dwarfs
and giants have been used to constrain the logistic curve fitted to
the discoveries. Figure \ref{rate_of_discovery} shows a sample of
fitted logistic curves, each curve being weighted by the likelihood
of its carrying capacity given by equation \ref{eq:likelihood}.

@@

The goal is to compute an expected oil production coming from future
discoveries, which requires to combine the steps described in sections
\ref{sub:Discoveries-modeled-as} and \ref{sub:Likelihood-function-for}.

The method described in section \ref{sub:Likelihood-function-for}
yields probabilities for the total number of fields (including the
not yet discovered fields) in each size bins (called a scenario).
However, this likelihood method does not give the time distribution
of future discoveries. The aim is to use the likelihood function to
generate scenarios with their respective occurrence probability.

For a given scenario, the carrying capacity $K$ (= total number of
fields) is given for each size class. This is useful, because it resolves
the instability in fitting the logistic curve to the number of discovered
fields. The time distribution of the discoveries is then given for
each size class by the fitted logistic curve.

@@

As the total parameter space is too large to be sampled entirely,
a Monte Carlo technique is applied to compute the expected production
with confidence interval from future discoveries. In a nutshell, the
algorithm works as follows:
\begin{enumerate}
\item Draw a scenario based on its probability.
\item Compute the time distribution of new discoveries by fitting a logistic
curve for each size class.
\item For each discovery generate a size and the resulting production curve
based on the size distribution and production curves of existing fields.
\item Superpose all the production curves.
\item Repeat and average over all drawn scenarios.
\end{enumerate}
The result is the expected production curve of future oil field discoveries.
The distribution of generated scenarios yields the confidence interval.

In order to account for the new fields discussed in section \ref{sub:Regular,-irregular-and},
which are already discovered but did not yet enter the decay phase,
the simulated production resulting from new discoveries has been shifted
to the date in the past where it matches the current production from
new fields.

@@

Based on the methodology described in section \ref{sub:Discovery-rate-of},
simulating future oil production was straightforward. For each country,
the existing fields were extrapolated and the future discoveries were
simulated. Figure \ref{no_uk_2013} shows the average of 1000 simulations.
For each country, the non-symmetric shape of the production's dynamics
contrary to what Hubbert would have predicted is immediately noticeable.

\begin{table}[H]
\caption{Remaining oil reserves until 2030 in barrels for the \textbf{Fit}
of the country's production, the Monte-Carlo \textbf{Model} and the
relative difference $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$
between the two.}

\centering %
\begin{tabular}{lllr}
 & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $31\%$\tabularnewline
U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $40\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

The results in table \ref{forecast-14} show a striking difference
between the fit and the Monte-Carlo model forecast. According to the
fit (extrapolation of aggregate production), Norway's future oil production
would decay much faster than in the Monte-Carlo case. The remaining
reserves estimated with the Monte-Carlo methodology are $30\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is higher
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the Monte-Carlo one is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$40\%$ compared to the latter.

Which of the two models is more thrust worthy? Clearly, the implications
in adopting one methodology over the other are significant. The only
way to answer this question is to back-test them. In other words:
\textquotedbl{}What would have each of the models predicted had they
been used in the past?\textquotedbl{}

@@

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the outcome of those tests. Comparing the forecast of both models
with the oil production of the subsequent 6 years, shows that the
predictive power of the Monte-Carlo model is better than a simple
extrapolation of aggregate production. The following table \ref{backtest-08}
summarizes the difference between the two approaches for the back-testing
period.

\begin{table}[H]
\caption{The fit and the Monte-Carlo model are used on the data set truncated
in 2008 and their forecast for the period 2008-2014 is compared to
the actual production.}

\centering %
\begin{tabular}{llll}
 & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
\midrule Norway & $133\cdot10^{6}$  & $108\cdot10^{6}$  & $130\cdot10^{6}$ \tabularnewline
U.K. & $79\cdot10^{6}$  & $66\cdot10^{6}$  & $75\cdot10^{9}$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}

The difference is not huge for the period from 2008 to 2014, but the
Monte-Carlo model nevertheless is noticeably closer to the actual
value. However, when considering the period after 2014, the difference
is huge as shown in table \ref{forecast-08}.

\begin{table}[H]
\caption{Remaining oil reserves forecasted for the period 2014-2030 when using
the data truncated in 2008.}

\centering %
\begin{tabular}{lllc}
 & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
\midrule Norway 08 & $25\cdot10^{6}$  & $171\cdot10^{6}$  & $-85\%$\tabularnewline
U.K. 08 & $25\cdot10^{6}$  & $91\cdot10^{9}$  & $-73\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-08}
\end{table}

>>>>

The idea behind the methodology is to model the future aggregate oil
production of a country by studying the production dynamics of its
individual constituents, the oil fields. The main benefit of this
approach, compared to working directly with aggregate production data,
is the possibility to forecast non-trivial oil production profiles
arising from the combination of all the individual field
dynamics. This ability reflects directly that the total
production is indeed the sum of the contributions of the individual
oil fields in production. Thus modelling at the level of each field
reflects more closely the reality and is likely to be over-performing
and more reliable, as we show below.
In order to implement our approach, one must be able 1) to \textbf{extend
the oil production of each individual field} into the future and 2)
to \textbf{extrapolate the rate of discoveries} of new oil fields.

@@

The first step to predicting the future oil production of a country
is to extrapolate the future production of existing fields and to
estimate the error on this extrapolation. The data of the fields developed
in the past shows a repeating asymmetric pattern. A good example is
the Oseberg field shown in figure \ref{regular}, with a quick ramp
up once the field is being developed, and then a peak or plateau before
the oil field production starts decaying. The decay can take many different shapes
and is governed by a variety of geological and economical factors.
The goal of the fitting procedure is to capture as much as possible the impact of
these different factors.

@@

To determine how well the extrapolation based on the stretched exponential
predicts the future production, a complete back-testing has been performed
on each field. A single back-test is made as follows:
\begin{itemize}
\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field is
truncated at a certain date in the past $T \in\{0,\,\ldots,\, N\}$, where
$T$ is the time counted in months since the production start of the field.
\item The extrapolation of the oil production rate
$f(t) := \frac{dP}{dt}$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{T}\}$.
\item The future production predicted by the extrapolation function $f(t)$ can be
compared to the actual production from the date $T$ in the past
up to the present $T_{f}=N$. The extrapolated total production can be
computed as 
\begin{equation}
P_{e}(T)=\int_{T}^{T_{f}}f(t)dt
\end{equation}
and the relative error is given by
\begin{equation}
e(T)=\frac{P_{e}(T)-\sum_{i=T}^{T_{f}}p_{i}}{P_{e}(T)},
\label{ryjryukoik}
\end{equation}
where both $P_{e}(T)$ and $e(T)$ are functions of the truncation time $T$.
\end{itemize}
Computing this back-test, for every month in the past since the field production
started decaying, yields a plot showing the evolution of the relative
error over time defined by expression (\ref{oesberg-error}). By construction, the relative
error will tend to zero as the truncation time $T$ approaches the
present. Nonetheless, it is a useful indicator for the stability of
the extrapolation. As can been seen for the Oseberg field in figure \ref{oesberg-error}, the relative
error on future production remained fairly stable during the past
decade.

From the complete back-test, we compute the average relative error
\begin{equation}
\bar{e}=\frac{1}{N}\sum_{i=0}^{N}e(i)
\end{equation}
of the extrapolation made on the future production. Assuming that the
relative errors are normally distributed around the average relative
error, the standard deviation on the average relative error is given
by 
\begin{equation}
\sigma_{e}=\sqrt{\frac{1}{N}\sum_{i=0}^{N}\left(e(i)-\bar{e}\right)^{2}}.\label{eq:field-risk}
\end{equation}
As the average relative error is often fairly constant, the extrapolation
was corrected by the average relative error, that is, if the extrapolation
consistently over-estimated the production by 10\% during the back-test,
the extrapolation was reduced by 10\%. This results in an extrapolated
production $p(t)$, including a 1$\sigma$ confidence interval, given by
\begin{equation}
p(t)=(1-\bar{e})\, f(t)\pm\sigma_{e}.
\label{eq:extrapolation}
\end{equation}
An example of such an extrapolation including a one standard deviation
range is shown in figure \ref{regular} for the Oseberg field.

@@

Once the individual fields have been extrapolated using
formula \ref{eq:extrapolation}, we compute the extrapolation of the oil production
for the whole country. While it is straightforward to sum the extrapolations
of the individual fields to obtain the expected production, some care
has to be taken with respect to the confidence interval of the production
at the country level.

As shown later in section \ref{sec:Results}, the same extrapolation
including a complete monthly back-test of total future production
has been performed at the country level and the resulting relative
error is much smaller than the average error observed on the individual
fields. To account for this observation, the assumption made is that
the relative error between individual fields is uncorrelated. 
While one could imagine that some inter-dependence could result from
a coordinated response of supply to a sharp increase/decrease of demand,  we have not 
observed this to be the case at a significant level. Therefore,
the fields can be considered as a portfolio of assets with a return
given by their extrapolation $p(t)$ (eq. \ref{eq:extrapolation})
and a risk given by $\sigma_{field}$ (eq. \ref{eq:field-risk}).
This means that the average standard deviation per field
at the country level from the extrapolated production can be computed as 
\begin{equation}
\sigma_{country}^{2}=\frac{1}{\#fields}\sum_{field\in fields}\sigma_{field}^{2}.
\end{equation}
Intuitively, this models well the fact that the uncorrelated errors
among fields will mostly cancel out.

@@

Knowing the future production rate of existing fields is not enough
as new fields will be discovered in the future. The model describing
the discovery rate of new fields should satisfy two fundamental observations.
\begin{enumerate}
\item The rate of new discoveries should tend to zero as time goes to infinity.
This is a consequence of the finiteness of the number of oil fields. 
\item The rate of new discoveries should depend on the size of the oil fields.
As of today, giant oil fields are discovered much less frequently
than dwarf oil fields.
\end{enumerate}

@@

A natural choice for such a model is a non-homogenous Poisson process.
The Poisson process is a process that generates independent events
at a rate $\lambda$. It is nonhomogeneous if the rate is time-dependent,
$\lambda\rightarrow\lambda(t)$. The standard way to measure $\lambda(t)$
is to find a functional form for $N(t)$, the statistical average
of the cumulative number of events (discoveries) up to time $t$.
Then, $\lambda(t)$ is simply a smoothed estimation 
of the observed rate $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
shows $N(t)$ for Norwegian fields classified according to their size in two classes, 
dwarfs and giant fields. 
The logistic curve is a good fit to the data (integral form of equation
\ref{logistic}). This implies that after an initial increase, the
rate of new discoveries reaches a peak followed by a decrease until
no more oil fields are to be found, consistent with our fundamental
observations. This same approach has already been successfully applied
by Forr\'o et al. \cite{Forro 2012} to estimate the number of daily active
users on Zynga.

As the discovery and production dynamics are not independent of the
field size, the fields have been split into two groups: dwarfs and
giants. Unfortunately, the two logistic curves thus obtained are highly
sensitive to the splitting size. This results from the major issue,
when fitting a logistic curve to data, that the carrying capacity
can not be determined if the data does not already exhibit the slowdown 
in growth towards the carrying capacity. However, it is
mentioned in the literature that often dwarf fields have already been
discovered a long time ago, but their production has been postponed
for economical reasons \cite[p. 378]{Lynch 2002}. Therefore, it is
expected that the large oil fields have mostly been found and produced,
and that future discoveries will mostly be made up of dwarf fields.
Consequently, the splitting size has been picked as small as possible
in order to maximize the number of giant fields but large enough to
avoid recent discoveries. Our definition is thus:
\begin{itemize}
\item \textbf{Dwarfs}: Fields which produced less than $50\cdot10^{6}$
barrels.
\item \textbf{Giants}: Fields which produced more than $50\cdot10^{6}$
barrels.
\end{itemize}
We note that this definition differs by a factor 10 from the more standard one, for which
oil fields with an ultimate recoverable resource of 0.5 billion barrels (Gb) 
or higher are classified as giants, while oil fields with smaller URR are considered to be dwarfs
\cite{H=0000F6=0000F6k 2008}.

The resulting plot shown in figure \ref{rate_of_discovery} pictures
the dynamics: giant oil fields have mostly been found while the discovery
process for dwarf fields is still ongoing. The logistic growth curve
fit to the giant discoveries is well constrained, however the fit
for the dwarfs is poorly constrained. As can be seen in figure \ref{rate_of_discovery},
the carrying capacity $K$ of the logistic growth model is not well
constrained by the available data. A large spectrum of values for
$K$ can lead to an equally good fit of the data. We have not taken
into account the possible effect that
newly discovered dwarf fields could become smaller and smaller 
until the new fields have a size that is so small to be economical to drill.
This effect would cap the carrying capacity at a value smaller than 
the one shown in figure  \ref{rate_of_discovery}. To address these issues, the method
described in the next section \ref{sub:Likelihood-function-for} has been used
to compute the probability of different carrying capacities.

@@

To overcome the poor constraint on the carrying capacity
$K$ obtained from the fitting procedure for dwarf fields, a method
already used by Smith \cite{Smith 1980} has been implemented. This
method makes the following two postulates:
\begin{enumerate}
\item ``The discovery of reservoirs in a petroleum play can be modeled
statistically as sampling without replacement from the underlying
population of reservoirs.''
\item ``The discovery of a particular reservoir from among the existing
population is random, with a probability of discovery being dependent 
on (proportional to) reservoir size.''
\end{enumerate}
The fields are split into $J$ size bins denoted $S_{1},\,\ldots,\, S_{J}$
occurring with frequency $n_{1},\,\ldots,\, n_{J}$. Each discovery
is considered as a step $i$ at which a field of size $I(i)\in\{S_{1},\,\ldots,\, S_{J}\}$
is found and $m_{ij}$ denotes the number of fields of size $j\in\{1,\,\ldots,\, J\}$
discovered before the $i^{th}$ step. Then, the probability that the
discovery at step $i$ is of size $j$ can be expressed as
\begin{equation}
P\left(I(i)=S_{j}\right)=\frac{\left(n_{j}-m_{ij}\right)\cdot S_{j}}{\sum_{k=1}^{J}\left(n_{k}-m_{ik}\right)\cdot S_{k}}.
\end{equation}
The likelihood $L$ for a complete sequence of $N$ discoveries $\{I(1),\,\ldots,\, I(N)\}$
can then be expressed as
\begin{equation}
L=\prod_{i=1}^{N}\frac{\left(n_{I(i)}-m_{iI(i)}\right)\cdot S_{I(i)}}{\sum_{j=1}^{J}\left(n_{j}-m_{ij}\right)\cdot S_{j}}.\label{eq:likelihood}
\end{equation}
The unknown parameters are the number of fields $n_{1},\,\ldots,\, n_{J}$,
whose likelihood can now be estimated based on the existing discoveries.
Using a brute force approach, the entire space of plausible values
for the variables $n_{1},\,\ldots,\, n_{J}$ has been sampled. The
values $n_{j}$ have been sampled between the number of existing fields
$m_{Nj}$ in the bin $j$ and up to a value $n_{j}^{upper}$, such
that the scenario with the largest likelihood according to equation
(\ref{eq:likelihood}) has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
the bin $j$. Subsequently, the likelihood of each scenario (value
of the tuple $n_{1},\,\ldots,\, n_{J}$) has been normalized such
that the total likelihood of all generated scenarios equals one.

For the analysis of the discoveries of the North Sea oil fields, the number
of size bins has been fixed to $J=2$ splitting between dwarfs (1)
and giants (2) as described in section \ref{sub:Discoveries-modeled-as}.

\begin{table}[h]
\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.
For Norway, our logistic fit suggests that up to two new giant fields could be
ultimately discovered. For the U.K., the prediction is more bleak, suggesting
that the most probable scenario is that no other giant field will be found.}
\centering %
\begin{tabular}{lllll}
 & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{likelihood-results}
\end{table}

The results shown in table \ref{likelihood-results} are coherent
with the intuitive expectation that discovering a new giant field
is unlikely and that future discoveries will mostly be made up of dwarf
fields. The likelihoods obtained for the carrying capacities of dwarfs
and giants have been used to constrain the logistic curve fitted to
the discoveries. Figure \ref{rate_of_discovery} shows a sample of
fitted logistic curves, each curve being weighted by the likelihood
of its carrying capacity given by equation (\ref{eq:likelihood}).

@@

We now compute an expected oil production coming from future
discoveries, which requires to combine the steps described in sections
\ref{sub:Discoveries-modeled-as} and \ref{sub:Likelihood-function-for}.

The method described in section \ref{sub:Likelihood-function-for}
yields probabilities for the total number of fields (including the
not yet discovered fields) in each size bins (called a scenario).
However, this likelihood method does not give the time distribution
of future discoveries. We propose to use the likelihood function to
generate scenarios with their respective occurrence probability.

For a given scenario, the carrying capacity $K$ (= total number of
fields) is given for each size class. This is useful to resolve
the instability in fitting the logistic curve to the number of discovered
fields. The time distribution of the discoveries is then given for
each size class by the fitted logistic curve.

@@

As the total parameter space is too large to be sampled entirely,
a Monte Carlo technique is applied to compute the expected production
with confidence interval from future discoveries. In a nutshell, the
algorithm works as follows:
\begin{enumerate}
\item Draw a scenario (total number of fields in each size bin) based on its probability according to the likelihood function (\ref{eq:likelihood}). This is done by generating a random number $r$ between 0 and 1, and computing the scenario which is mapped to $r$ by the cumulative distribution function of all scenarios.
\item Compute the time distribution of new discoveries by fitting a logistic
curve for each size class.
\item For each discovery, generate a size and the resulting production curve
based on the size distribution and production curves of existing fields.
\item Superpose all the production curves.
\item Repeat and average over all drawn scenarios.
\end{enumerate}
The result is the expected production curve of future oil field discoveries.
The distribution of generated scenarios yields the confidence interval.

Last but not least, it has to be defined how the expected production from these future oil field discoveries is added to the extrapolated production from existing fields. The start of the simulation of new discoveries does not match up with the date of the latest production data. The reason being that the new fields (defined in section \ref{sub:Regular,-irregular-and}),
which are already discovered but did not yet enter the decay phase, are not taken into account in the simulation as their final size is not known. The only meaningful way of treating the new fields is to consider them as a discovery. Therefore, the starting point in time of the simulated production resulting from new discoveries has been choosen as the date in the past where it matches the current production from
new fields. In order words, the extrapolated production from regular and irregular fields added to the production from simulated future discoveries (which is shifted into the past as to match the production from new fields) must be equal to the current (latest available data) total production from all fields.

@@

Based on the methodology described in section \ref{sub:Discovery-rate-of},
simulating future oil production was straightforward. For each country,
the existing oil field productions were extrapolated and the future discoveries were
simulated. Figure \ref{no_uk_2013} shows the average of 1000 simulations.
For each country, the non-symmetric shape of the production dynamics, which
contradicts the prediction based on Hubbert's standard approach,
is immediately noticeable.

\begin{table}[H]
\caption{Remaining oil reserves until 2030 in barrels 
predicted by the extrapolation of the  \textbf{Fit}
of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
The relative difference between these two predictions 
is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$.}
\centering %
\begin{tabular}{lllr}
 & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $31\%$\tabularnewline
U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $40\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

The results in table \ref{forecast-14} show a striking difference
between the extrapolation of the fit and the Monte-Carlo model forecast. According to the
fit (extrapolation of aggregate production), Norway's future oil production
would decay much faster than in the Monte-Carlo case. The remaining
reserves estimated with the Monte-Carlo methodology are $30\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue for us to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the Monte-Carlo one is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$40\%$ compared to the latter.

Which of the two models is more trustworthy? Clearly, the implications
in adopting one methodology over the other are significant. The only
way to answer this question is to back-test them. In other words:
\textquotedbl{}What would each of the models have predicted, had they
been used in the past?\textquotedbl{} The next section addresses
this question and presents the validation step of our approach.

@@

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the results of these tests. Comparing the forecast of both models,
with the oil production of the subsequent 6 years, shows that the
predictive power of the Monte-Carlo model is significantly better than a simple
extrapolation of the aggregate past production. The Monte-Carlo model 
is found essentially right on target, while the extrapolation of past production
(``fit'') is under-estimating the realised production by 19\% and 16\% respectively
for Norway and the U.K. The following table \ref{backtest-08}
summarizes the difference between the two approaches for the back-testing
period.

\begin{table}[H]
\caption{Extrapolation of past oil production (``fit'') and prediction
using the Monte-Carlo model are used on the data set truncated
in 2008. Their forecast for the period 2008-2014 is compared to
the actual realised production.}
\centering %
\begin{tabular}{llll}
 & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
\midrule Norway & $133\cdot10^{6}$  & $108\cdot10^{6}$  & $130\cdot10^{6}$ \tabularnewline
U.K. & $79\cdot10^{6}$  & $66\cdot10^{6}$  & $75\cdot10^{9}$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}

While the error of the extrapolation of the past oil production method (``fit'') is not dramatic
over these six years from 2008 to 2014, the difference between it and the Monte-Carlo
approach becomes huge from 2014 to 2030, as shown in table \ref{forecast-08}.
